The Courtroom Challenge—A Conjecture on the Unavoidable Cost of LLM Verification

### 2025-06-20 paper/main.tex
feat: add LaTeX sources and figures under paper/

% =========================================================
\begin{document}
\maketitle
\begin{abstract}
  Large language models hallucinate; ad-hoc patches conceal the true cost of verification. 
  We therefore pose a falsifiable ultimatum: either automate validation, prove it unnecessary, or pay the explicit human-in-the-loop cost.  
  Conjecture 1 (the Courtroom Challenge). 
  For any unbounded open-world text stream, a system that lacks both (i) a persistent precedent store and (ii) a friction-priced validation loop will accumulate unbounded prequential regret.
  We formalise the state machine, cost ledger, and impossibility triangle that make this conjecture testable, and invite the community to break it.
  A 100-query toy stream illustrates how the ledger is logged\footnote{Baseline script will appear at \url{https://github.com/johnrepsys/courtroom-model-lean} with v2.}; no attempt is made here to optimise the validator.
  In addition, we present the Courtroom Model, a wrapper architecture that subjects every LLM claim to a perpetual cycle of precedent search, automated semantic checks, and selective escalation to an external judge—human or automated.
  Each new output is validated against a persistent vector memory, while conflicting evidence triggers the re-examination of prior “case law.”
  The challenge stands: thrive in an open world without paying the validation bill—or show why the bill is inescapable.
\end{abstract}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.9\linewidth]{courtroom}%
      % no .pdf extension needed if LaTeX finds courtroom.pdf
  \caption{One step of the Courtroom loop.%
           \ Solid arrows: primary query flow.%
           \ Dashed arrows: helper calls and cost logging.%
           \ The cost\,bus aggregates all resource writes into a single ledger.
           \ On \textsc{DEFER} the round is skipped (no cost, no memory write).}
  \label{fig:courtroom}
\end{figure}

@@

>>>>

% =========================================================
\begin{document}
\maketitle
\begin{abstract}
  Large language models hallucinate; ad-hoc patches conceal the true cost of verification. 
  We therefore pose a falsifiable ultimatum: either automate validation, prove it unnecessary, or pay the explicit human-in-the-loop cost.  
  Conjecture 1 (the Courtroom Challenge). 
  For any unbounded open-world text stream, a system that lacks both (i) a persistent precedent store and (ii) a friction-priced validation loop will accumulate unbounded prequential regret.
  We formalise the state machine, cost ledger, and impossibility triangle that make this conjecture testable, and invite the community to break it.
  A 100-query toy stream illustrates how the ledger is logged\footnote{Baseline script will appear at \url{https://github.com/johnrepsys/courtroom-model} with v2.}; no attempt is made here to optimise the validator.
  In addition, we present the Courtroom Model, a wrapper architecture that subjects every LLM claim to a perpetual cycle of precedent search, automated semantic checks, and selective escalation to an external judge—human or automated.
  Each new output is validated against a persistent vector memory, while conflicting evidence triggers the re-examination of prior “case law.”
  The challenge stands: thrive in an open world without paying the validation bill—or show why the bill is inescapable.
\end{abstract}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.9\linewidth]{courtroom}%
      % no .pdf extension needed if LaTeX finds courtroom.pdf
  \caption{One step of the Courtroom loop.%
           \ Solid arrows: primary query flow.%
           \ Dashed arrows: helper calls and cost logging.%
           \ The cost\,bus aggregates all resource writes into a single ledger.
           \ On \textsc{DEFER} the round is skipped (no cost, no memory write).
           \ If C cannot supply an authoritative value—because the fact is genuinely unknown—the 
           \ query transitions to a terminal UNRESOLVED state; no memory write occurs.}
  \label{fig:courtroom}
\end{figure}

@@

\paragraph{Provisionality of verdicts.}
Even the external Judge’s verdicts are provisional: they may be rescinded
if subsequent audits or new primary filings contradict earlier
conclusions.  Accordingly, Judge-validated precedents can be re-queued,
incur additional \(\WriteCost\), and overwrite prior entries via the usual
supersession mechanism.
